---
title: "Categorical Variable Price Distributions"
author: "Graeme Keleher"
date: "July 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Purpose
The point of this notebook is to plot categorial features to see if they have linear structure when converted to numerical 


## Load Tidyverse Suite of Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)  #Data wrangling and plotting
library(scales)     #Formatting axis
```

## Import Training Data
```{r message=FALSE, warning=FALSE}
train <- read_csv("train.csv")
```


#LandContour

```{r, echo = FALSE}

train %>%
  group_by(LandContour)%>%
  summarise(class_total = n())%>%
  ggplot(aes(LandContour, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by LandContour",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(LandContour, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs LandContour",
       y = "Sale Price",
       x = "LandContour")+ 
  scale_y_continuous(labels = comma)

```


#BsmtCond

```{r, echo = FALSE}

train %>%
  group_by(BsmtCond)%>%
  summarise(class_total = n())%>%
  ggplot(aes(BsmtCond, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by BsmtCond",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(BsmtCond, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs BsmtCond",
       y = "Sale Price",
       x = "BsmtCond")+ 
  scale_y_continuous(labels = comma)

```


#BsmtExposure

```{r, echo = FALSE}

train %>%
  group_by(BsmtExposure)%>%
  summarise(class_total = n())%>%
  ggplot(aes(BsmtExposure, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by BsmtExposure",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(BsmtExposure, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs BsmtExposure",
       y = "Sale Price",
       x = "BsmtExposure")+ 
  scale_y_continuous(labels = comma)

```


#BsmtFinType1

```{r, echo = FALSE}

train %>%
  group_by(BsmtFinType1)%>%
  summarise(class_total = n())%>%
  ggplot(aes(BsmtFinType1, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by BsmtFinType1",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(BsmtFinType1, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs BsmtFinType1",
       y = "Sale Price",
       x = "BsmtFinType1")+ 
  scale_y_continuous(labels = comma)

```


#BsmtFinType2

```{r, echo = FALSE}

train %>%
  group_by(BsmtFinType2)%>%
  summarise(class_total = n())%>%
  ggplot(aes(BsmtFinType2, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by BsmtFinType2",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(BsmtFinType2, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs BsmtFinType2",
       y = "Sale Price",
       x = "BsmtFinType2")+ 
  scale_y_continuous(labels = comma)

```



#BsmtQual

```{r, echo = FALSE}

train %>%
  group_by(BsmtQual)%>%
  summarise(class_total = n())%>%
  ggplot(aes(BsmtQual, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by BsmtQual",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(BsmtQual, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price VsBsmtQual",
       y = "Sale Price",
       x = "BsmtQual")+ 
  scale_y_continuous(labels = comma)

```


#ExterCond

```{r, echo = FALSE}

train %>%
  group_by(ExterCond)%>%
  summarise(class_total = n())%>%
  ggplot(aes(ExterCond, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by ExterCond",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(ExterCond, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs ExterCond",
       y = "Sale Price",
       x = "ExterCond")+ 
  scale_y_continuous(labels = comma)

```


#ExterQual

```{r, echo = FALSE}

train %>%
  group_by(ExterQual)%>%
  summarise(class_total = n())%>%
  ggplot(aes(ExterQual, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by ExterQual",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(ExterQual, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs ExterQual",
       y = "Sale Price",
       x = "ExterQual")+ 
  scale_y_continuous(labels = comma)

```


#FireplaceQu

```{r, echo = FALSE}

train %>%
  group_by(FireplaceQu)%>%
  summarise(class_total = n())%>%
  ggplot(aes(FireplaceQu, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by FireplaceQu",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(FireplaceQu, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs FireplaceQu",
       y = "Sale Price",
       x = "FireplaceQu")+ 
  scale_y_continuous(labels = comma)

```



#GarageCond

```{r, echo = FALSE}

train %>%
  group_by(GarageCond)%>%
  summarise(class_total = n())%>%
  ggplot(aes(GarageCond, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by GarageCond",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(GarageCond, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs GarageCond",
       y = "Sale Price",
       x = "GarageCond")+ 
  scale_y_continuous(labels = comma)

```


#GarageQual

```{r, echo = FALSE}

train %>%
  group_by(GarageQual)%>%
  summarise(class_total = n())%>%
  ggplot(aes(GarageQual, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by GarageQual",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(GarageQual, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs GarageQual",
       y = "Sale Price",
       x = "GarageQual")+ 
  scale_y_continuous(labels = comma)

```


#HeatingQC

```{r, echo = FALSE}

train %>%
  group_by(HeatingQC)%>%
  summarise(class_total = n())%>%
  ggplot(aes(HeatingQC, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by HeatingQC",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(HeatingQC, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs HeatingQC",
       y = "Sale Price",
       x = "HeatingQC")+ 
  scale_y_continuous(labels = comma)

```


#KitchenQual

```{r, echo = FALSE}

train %>%
  group_by(KitchenQual)%>%
  summarise(class_total = n())%>%
  ggplot(aes(KitchenQual, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by KitchenQual",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

g<- train %>%
  ggplot(aes(reorder(KitchenQual, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs KitchenQual",
       y = "Sale Price",
       x = "KitchenQual")+ 
  scale_y_continuous(labels = comma)

ggsave("plots/Kitchen_quality.jpeg",plot = g,dpi = 320, width = 8, height = 4)

```


#PoolQC

```{r, echo = FALSE}

train %>%
  group_by(PoolQC)%>%
  summarise(class_total = n())%>%
  ggplot(aes(PoolQC, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by PoolQC",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(PoolQC, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs PoolQC",
       y = "Sale Price",
       x = "PoolQC")+ 
  scale_y_continuous(labels = comma)

```



#LandContour

```{r, echo = FALSE}

train %>%
  group_by(LandContour)%>%
  summarise(class_total = n())%>%
  ggplot(aes(LandContour, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by LandContour",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(LandContour, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs LandContour",
       y = "Sale Price",
       x = "LandContour")+ 
  scale_y_continuous(labels = comma)

```

#Utilities

```{r, echo = FALSE}

train %>%
  group_by(Utilities)%>%
  summarise(class_total = n())%>%
  ggplot(aes(Utilities, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Utilities",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(Utilities, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Utilities",
       y = "Sale Price",
       x = "Utilities")+ 
  scale_y_continuous(labels = comma)

```

#LotShape

```{r, echo = FALSE}

train %>%
  group_by(LotShape)%>%
  summarise(class_total = n())%>%
  ggplot(aes(LotShape, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by LotShape",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(LotShape, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs LotShape",
       y = "Sale Price",
       x = "LotShape")+ 
  scale_y_continuous(labels = comma)

```


#PavedDrive

```{r, echo = FALSE}

train %>%
  group_by(PavedDrive)%>%
  summarise(class_total = n())%>%
  ggplot(aes(PavedDrive, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by PavedDrive",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(PavedDrive, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs PavedDrive",
       y = "Sale Price",
       x = "PavedDrive")+ 
  scale_y_continuous(labels = comma)

```




#Functional

```{r, echo = FALSE}

train %>%
  group_by(Functional)%>%
  summarise(class_total = n())%>%
  ggplot(aes(Functional, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Functional",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(Functional, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Functional",
       y = "Sale Price",
       x = "Functional")+ 
  scale_y_continuous(labels = comma)

```


#LandSlope

```{r, echo = FALSE}

train %>%
  group_by(LandSlope)%>%
  summarise(class_total = n())%>%
  ggplot(aes(LandSlope, class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by LandSlope",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(reorder(LandSlope, SalePrice, mean), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs LandSlope",
       y = "Sale Price",
       x = "LandSlope")+ 
  scale_y_continuous(labels = comma)

```