---
title: "Numerical or Categorical?"
author: "Graeme Keleher"
date: "July 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Purpose
The point of this notebook is to plot numerical features and determine if they should actually be categorical


## Load Tidyverse Suite of Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)  #Data wrangling and plotting
library(scales)     #Formatting axis
```

## Import Training Data
```{r message=FALSE, warning=FALSE}
train <- read_csv("train.csv")
```

#GarageCars
Should be transformed to categorical

```{r, echo = FALSE}

train %>%
  group_by(GarageCars)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(GarageCars), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Garage Size",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

g <- train %>%
  ggplot(aes(as.factor(GarageCars), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs GarageCars",
       y = "Sale Price",
       x = "GarageCars")+ 
  scale_y_continuous(labels = comma)


ggsave("plots/cars_in_garage.jpeg",plot = g,dpi = 320, width = 8, height = 4)
```

## OverallCond
POSSIBLY turn to categorical, or maybe condense 5 and above to 5?


```{r, echo=FALSE}
train %>%
  group_by(OverallCond)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(OverallCond), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Overall Condition",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(OverallCond), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Overall Condition",
       y = "Sale Price",
       x = "Overall COndition")+ 
  scale_y_continuous(labels = comma)

```




## MSSubClass
Should be transformed to categorical


```{r, echo=FALSE}
train$MSSubClass <- as.factor(train$MSSubClass)

levels(train$MSSubClass) <- list("1-STORY 1946 & NEWER ALL STYLES" = "20",
                  "1-STORY 1945 & OLDER" = "30",
                  "1-STORY W/FINISHED ATTIC ALL AGES" = "40",
                  "1-1/2 STORY - UNFINISHED ALL AGES" = "45",
                  "1-1/2 STORY FINISHED ALL AGES" = "50",
                  "2-STORY 1946 & NEWER" = "60",
                  "2-STORY 1945 & OLDER" = "70",
                  "2-1/2 STORY ALL AGES" = "75",
                  "SPLIT OR MULTI-LEVEL" = "80",
                  "SPLIT FOYER" = "85",
                  "DUPLEX - ALL STYLES AND AGES" = "90",
                  "1-STORY PUD (Planned Unit Development) - \n1946 & NEWER" = "120",
                  "1-1/2 STORY PUD - ALL AGES" = "150",
                  "2-STORY PUD - 1946 & NEWER" = "160",
                  "PUD - MULTILEVEL - INCL SPLIT \nLEV/FOYER" = "180",
                  "2 FAMILY CONVERSION - ALL STYLES \nAND AGES" = "190")



train %>%
  group_by(MSSubClass)%>%
  summarise(class_total = n())%>%
  ggplot(aes(reorder(MSSubClass,class_total), class_total))+
  geom_bar(stat = "identity")+
  coord_flip()+
  labs(title = "Number of Sales by Dwelling Class (MSSubClass)",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)



train %>%
  ggplot(aes(reorder(MSSubClass, SalePrice), SalePrice))+
  geom_boxplot()+
  coord_flip()+
  labs(title = "Sale Price by Dwelling Class (MSSubClass)",
       y = "Sale Price",
       x = "")+ 
  scale_y_continuous(labels = comma)
```


## OverallQual
Don't transform to categorical, possible transformation needed


```{r, echo=FALSE}
train %>%
  group_by(OverallQual)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(OverallQual), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Overall Quality",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(OverallQual), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs OverallQual (1-10)",
       y = "Sale Price",
       x = "OverallQual (1-10)")+ 
  scale_y_continuous(labels = comma)

```

## BsmtFullBath
Perhaps merge 1-3?


```{r, echo=FALSE}
train %>%
  group_by(BsmtFullBath)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(BsmtFullBath), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of full baths in basement",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(BsmtFullBath), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Full baths in basement",
       y = "Sale Price",
       x = "Full Baths in Basement")+ 
  scale_y_continuous(labels = comma)

```

## BsmtHalfBath
Can probably just get rid of this feature...


```{r, echo=FALSE}
train %>%
  group_by(BsmtHalfBath)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(BsmtHalfBath), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of half baths in basement",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(BsmtHalfBath), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Half baths in basement",
       y = "Sale Price",
       x = "Half Baths in Basement")+ 
  scale_y_continuous(labels = comma)

```

## FullBath
combine 0 with 1


```{r, echo=FALSE}
train %>%
  group_by(FullBath)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(FullBath), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of full baths",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(FullBath), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Full baths",
       y = "Sale Price",
       x = "Full Baths")+ 
  scale_y_continuous(labels = comma)

```


## HalfBath
combine 1 and 2?


```{r, echo=FALSE}
train %>%
  group_by(HalfBath)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(HalfBath), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of half baths",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(HalfBath), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Half baths",
       y = "Sale Price",
       x = "Half Baths")+ 
  scale_y_continuous(labels = comma)

```


## TotalBath
Combine 0 and 1


```{r, echo=FALSE}
train %>%
  mutate(TotalBath = FullBath + HalfBath)%>%
  group_by(TotalBath)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(TotalBath), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of total baths",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  mutate(TotalBath = FullBath + HalfBath)%>%
  ggplot(aes(as.factor(TotalBath), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Total baths",
       y = "Sale Price",
       x = "Total Baths")+ 
  scale_y_continuous(labels = comma)

```



## BedroomAbvGr
can we condense to 1-4 bedrooms!?


```{r, echo=FALSE}
train %>%
  group_by(BedroomAbvGr)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(BedroomAbvGr), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of Bedrooms Above Ground",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(BedroomAbvGr), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Bedrooms Above Ground",
       y = "Sale Price",
       x = "Bedrooms Above Ground")+ 
  scale_y_continuous(labels = comma)

```

## KitchenAbvGr
condense to 1 & 2!?


```{r, echo=FALSE}
train %>%
  group_by(KitchenAbvGr)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(KitchenAbvGr), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of Kitchen Above Ground",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(KitchenAbvGr), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Kitchens Above Ground",
       y = "Sale Price",
       x = "Kitchen Above Ground")+ 
  scale_y_continuous(labels = comma)

```

## Fireplaces
Don't convert to categorical


```{r, echo=FALSE}
train %>%
  group_by(Fireplaces)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(Fireplaces), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of Fireplaces",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(Fireplaces), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Fireplaces",
       y = "Sale Price",
       x = "Fireplaces")+ 
  scale_y_continuous(labels = comma)

```


## KitchenAbvGr
Don't convert to categorical


```{r, echo=FALSE}
train %>%
  group_by(KitchenAbvGr)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(KitchenAbvGr), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of Kitchens Above GRound",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(KitchenAbvGr), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Kitchens Above GRound",
       y = "Sale Price",
       x = "Kitchens Above Ground")+ 
  scale_y_continuous(labels = comma)

```

## TotRmsAbvGrd
May be worth merging everything above 11 into 11


```{r, echo=FALSE}
train %>%
  group_by(TotRmsAbvGrd)%>%
  summarise(class_total = n())%>%
  ggplot(aes(as.factor(TotRmsAbvGrd), class_total))+
  geom_bar(stat = "identity")+
  labs(title = "Number of Sales by Number of Rooms Above GRound",
       y = "Number of Sales",
       x = "")+
  geom_text(aes(label=class_total),  size=3.5, nudge_y = 40)

train %>%
  ggplot(aes(as.factor(TotRmsAbvGrd), SalePrice))+
  geom_boxplot()+
  labs(title = "Sale Price Vs Number of Rooms Above GRound",
       y = "Sale Price",
       x = "Rooms Above Ground")+ 
  scale_y_continuous(labels = comma)

```